### 2017-09-16 manuscript/online-appendix.Rmd

---
title:  |
        | **Data Codebook and Appendix for**
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author:
  - "Danilo Freire^[PhD Candidate, Department of Political Economy, King's College London. Email:   [danilofreire@gmail.com](mailto:danilofreire@gmail.com).] &nbsp; John Meadowcroft^[Senior Lecturer in Public Policy, Department of Political Economy, King's College London. Email:  [john.meadowcroft@kcl.ac.uk](mailto:john.meadowcroft@kcl.ac.uk).] &nbsp; David Skarbek^[Associate Professor, Department of Political Science, Brown University. Email: [david_skarbek@brown.edu](david_skarbek@brown.edu). Corresponding author.] &nbsp; Eugenia Guerrero^[Software Developer, Tempo UK.]"
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: How did the Pinochet regime use deaths and disappearances after the violent military coup in September 1973? The coup resulted in more than three decades of violence as General Pinochet's government fought against left-wing movements for the sake of the 'national wellbeing.' While these efforts were significant, they were not uniformly distributed spatially or over time during his regime. This paper provides new data on observed variation of regime-based violence. It codes and compiles data from two volumes of the Report of the Chilean National Commission on Truth and Reconciliation, a systematic investigation of the Pinochet regime published in 1991. The data documents the initially high-level and then steep decline of these atrocities, the extent to which victims were targeted, and other demographic information about the victims. Our data also includes geographic information system data that allows researchers to map the spatial variation of atrocities over time.
keywords: authoritarianism, Chile, human rights, Pinochet, truth commission
fontsize: 12pt
monospace-url: no
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: no
    keep_tex: no
    template: template.latex
    latex_engine: pdflatex
---

\newpage

# Introduction

\setlength{\parindent}{0pt}

This document contains the codebook, data and code necessary to replicate the results in "Deaths and Disappearances in the Pinochet Regime" by @freire2017pinochet-regime-data. The `dataset.csv` file includes the complete dataset. The original Excel file, also included in the paper [GitHub repository](https://github.com/danilofreire/pinochet-victims-data), has been pre-processed to reduce compatibility problems with other statistical software such as R or Stata. All data pre-processing steps are described below. The analyses were performed in the R statistical language, version 3.4.1 [@rstats2017].

# Background

The violent military coup in Chile in September 1973 resulted in over three decades of sustained violence, deaths, and forced disappearances as General Pinochet’s government fought against left-wing movements for the sake of the 'national wellbeing'. The fall of Pinochet’s regime and the traumatic consequences of his authoritarian government left Chile deeply divided. In order to move forward and avoid a repetition of the past, the Chilean government highlighted the need of an impartial and comprehensive report detailing the extent of the deaths and disappearances.

By compiling and analyzing the data from both volumes of the Report of the Chilean National Commission on Truth and Reconciliation, we seek to understand the logic and dynamics of the violence, deaths, and disappearances under Pinochet’s Chile.

# Sources

Report of the Chilean National Commission on Truth and Reconciliation, Vol. 1.

Report of the Chilean National Commission on Truth and Reconciliation, Vol. 2.

\newpage

# Variables

**Region**. Report of the Chilean National Commission on Truth and Reconciliation, Vol. 1.

**Case number**. Variable label: “Case number in the dataset”.

**Sub-case number**. Variable label: “Sub-case number in the dataset”.

**Merged ID**. Variable label: “Identification number in the dataset”. The variable ‘Merged ID’ was created so as to have a complete row of ID numbers recognized by the mapping software. The previous two variables label cases by number (e.g. 1098) and subdivide the cases within that number (e.g. 1098 A, 1098 B).

**Page**. Variable label: “Page related to the identification number in the dataset”.

**Start Date**. Variable label: “Start date of incident”.

**End Date**. Variable label: “End date of incident”. This variable attempts to pinpoint when the individual died, was found, or was last presumed to be seen.

**LastName**. Variable label: “Last name(s) of individual”.

**FirstName**. Variable label: “First name(s) of individual”.

**Minor**. Variable label: “Minor or not”. Categories:

* 0 - Adult
* 1 - Minor
* NA - Unspecified

**Age**. Variable label: “Age of individual”. Range: from 2 to 85 years old.

**Female**. Variable label: “Female”. Categories:

* 0 - Male
* 1 - Female
* NA - Unknown

**Nationality**. Variable label: “Country of nationality label”. Categories:

* Argentinean
* Austrian
* Bolivian
* Brazilian
* Chilean
* Dual British and Chilean
* Dual French and Chilean
* Dual Swiss and Chilean
* Ecuadorian
* French
* Mexican
* Spanish
* Uruguayan
* Vietnamese

**Occupation**. Variable label: “Identification number of individual’s occupation”. Categories:

* 1 - White collar
* 2 - Blue collar
* 3 - School student
* 4 - University student
* 5 - Military
* 6 - Non-military government
* 7 - Unemployed
* 8 - Unknown
* 9 - Ex-military
* 10 - Housewife

**Occupation (string)**. Variable label: “Description of individual’s occupation”.

\newpage

**Victim Affiliation** Variable label: “Individual’s political affiliation”. Categories:

* 1 - _Opposition (Allende/Left)_: this category encompasses individuals loyal to the ex-president Allende, as well as individuals who were affiliated or leaders of left-wing groups
* 2 - _Regime (Pinochet)_: this category encompasses individuals who belonged to the Pinochet
government.
* 3 - _Unknown_: this category refers to individuals whose political affiliation was unknown or
unspecified.
* 4 - _None_: this category pertains to individuals whom the report states had no known political
affiliation.

**Victim Affiliation (string)**. Variable label: “Description of individual’s political affiliation” This complements the previous variable by giving as much information as possible on the victims’ political affiliations. Categories:

* _Opposition_
    - Communist
        - Young Communist
    - Christian Democrat/Christian Left
    - FER-MIR
    - MAPU (Movimiento de Acción Popular Unitaria)
    - MIR (Revolutionary Left Movement)
    - National Party
    - Pro-Democracy party (PPD)
    - Revolutionary Radical Party
        - Revolutionary Radical Youth Party
    - Socialist
        - Young Socialist
* _Regime_. This variable encompasses individuals who belonged to the Pinochet government.

\newpage

**Violence**. Categories:

* 1 - Killed: cases where the Commission signals a definite and known death.
* 2 - Suicide
* 3 - Disappearance: cases where the victims are presumed to be dead at the hands of government
agents. Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report.
* 4 - Disappearance, information of death: cases which the Commission has information that signals that “the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims’ mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion” (p.44).
* 5 - Unresolved: insufficient information or evidence

**Method**. Categories:

* 1 - Gun
* 2 - Knife
* 3 - Unknown
* 4 - Beaten
* 5 - Hung
* 6 - Denied medical treatment
* 7 - Acute loss of blood
* 8 - Asphyxiated
* 9 - Intentional car crash
* 10 - Cardio-respiratory arrest
* 11 - Torture
* 12 - Jumped
* 13 - Immersion
* 14 - Poisoned
* 15 - Bomb
* 16 - Burned
* 17 - Tear gas
* 18 - Electrocuted

**Interrogation**. Variable label: “Individual was interrogated”. Categories:

* 1 - Yes
* 2 - No
* 3 - Unknown

**Torture**. Variable label: “Individual was tortured”. Categories:

* 1 - Yes
* 2 - No
* 3 - Unknown

“For the purposes of this Convention, torture means any act by which severe pain or suffering, whether physical or mental, is intentionally inflicted on a person for such purposes as obtaining from him or a third person information or confession, punishing him for an act he or a third person has committed or is suspected of having committed, or intimidating or coercing him or a third person, or for any reason based on discrimination of any kind, when such pain or suffering is inflicted by or at the instigation of, or with the consent or acquiescence of a public official or other person acting in an official capacity. It does not include pain or suffering arising only from, inherent in or incidental to, lawful sanctions” (pp. 38-39).

“These agencies – the CNI, the DICOMAR, and the group called COVEMA – practiced torture. CNI’s use of torture was systematic but more selective than that of the DINA which tortured practically everyone who passed through some of its secret facilities. The main torture methods continued to be the use of electricity, especially on the sensitive parts of the body, all kinds of beatings, and plunging the person’s head down into the water to the brink of asphyxiation, and then doing it again/ There are also indications that DICOMAR and COVEMA practiced torture, even to the point of death, as this Commission has verified and as will be noted in the case material.” (p. 644)

**Mistreatment**. Variable label: “Individual was mistreated”.

**Mistreatment means**. Categories:

* 1 - Yes
* 2 - No
* 3 - Unknown

**Targeted**. Variable label: “Individual was targeted”. Categories:

* 1 - Denounced
* 2 - Summoned
* 3 - Listed
* 4 - Voluntarily reported
* 5 - At home
* 6 - Unspecified

**Press**. Variable label: “Individual’s disappearance or death was justified in the press”. Accounts of an individual’s disappearance or death often appeared in the press, but the Truth Report refutes many of these accounts on the basis that it covers up or legitimizes the violence stemming from Pinochet’s regime. For example, fictional gun battles were fabricated, legitimizing the arbitrary and often unprovoked deaths of those rebelling against the government. Categories:

* 1 - Yes
* 2 - Unspecified

**War tribunal**. Variable label: “Individual was subjected to a war tribunal”. Categories:

* 1 - Yes
* 2 - Unspecified

**Previous arrests**. Variable label: “Individual was previously arrested”. Categories:

* 0 - Reported to the authorities and was allowed to go free (no arrest)
* 1 - One arrest
* 2 - Two arrests
* 3 - Three arrests
* 4 - Four arrests

**Perpetrator affiliation**. Variable label: “Perpetrator’s political affiliation”. Categories:

* 1 - Opposition (Allende/Left): “private citizens who committed terrorist acts or other kinds of politically motivated attacks” (p.43)
* 2 - Regime (Pinochet):
* 3 - Unknown
* 4 - None

    - _‘Opposition’_ (1):
    - _‘Government agents’_ (2):
        - “In this report the Commission is offering as much information as it could obtain about actions committed by government agents [...]: when such information is available the report names the branch or branches said to have participated, and specifically the regiment, base, police precinct, garrison, or group from which the official forces came” (p.43)
        - “When the Commission was unable to obtain such information but did come to the conviction that the person was killed by, or disappeared in the hands of, government agents, it has stated so” (p.43)
    - When the perpetrator is _‘unknown’_ (3): “The Commission has not presumed that government agents were involved in the death of individuals, even when it is clear that they were killed by firearms and when there is every reason to believe that the motivation was political, unless there are some grounds for that judgment” (p.43). See also “References to private citizens” on p. 44
    - “When the Commission here refers to the perpetrators as private citizens acting for political reasons or pretexts, it does not always mean that these were people who were opposed to the military government. In some cases the political motives of such private citizens were quite the contrary, that is, they supported the government. In some of those cases, which will be narrated in the chapter on those killed during protest actions, the Commission does not rule out the possibility that such private citizens were really security agents in civilian dress”

**Perpetrator (string)**. Variable label: “Description of perpetrator’s political affiliation”.

**Locations**. Place. Categories:

* 1 - Home
* 2 - Work
* 3 - University
* 4 - In custody
* 5 - In public
* 6 - In hospital
* 7 - Unknown

**Location (string)**. Variable label: “Location where victim was spotted/reported to be seen”.

**Latitude**. Variable label: “Latitudinal coordinates pertaining to the string ‘location’ variable”.

**Longitude**. Variable label: “Longitudinal coordinates pertaining to the string ‘location’ variable”.

**Exactitude of coordinates**. Variable label: “Exactitude of coordinates pertaining to the string ‘location’ variable”. Categories:

* 0 - Exact
* 1 - Area/neighborhood level
* 2 - City/town level

_The determination of coordinates_. The compilation of the ‘Location’ variable was based completely on information given in the Truth Report. However, since this information was in a string format (e.g. intersection of Calle Grecia and Avenida Rosa), creating a new variable incorporating each location’s latitude and longitude was necessary to pursue further analysis of the trends in deaths and disappearances. The format chosen was decimal coordinates.

_Sources used for the determination of coordinates_

Three main websites were used to find the decimal coordinates pertaining to the string ‘Location’ variable:

* 1 - Latitude: http://latitude.to/. Latitude was mostly used for sourcing more inexact coordinates, such as the coordinates for cities or towns. It was also used to convert non-decimal coordinates into the decimal format.
* 2 - Wikimapia: http://wikimapia.org/. Wikimapia was mostly used for locating shantytowns, neighborhoods, military bases, or police stations that were not found on Google maps. Its strength is that it is an editable map, meaning that individual users can collaborate so as to produce a more comprehensive map (filled with local knowledge) than other standard maps such as Google maps.
* 3 - Google maps: http://maps.google.co.uk/. Google maps was useful for finding intersections of roads or precise addresses, as well as for searching for police stations, some military bases, and other governmental buildings.

\newpage

# Data Pre-Processing

## Load required packages
```{r, message=FALSE,warning=FALSE}
if (!require("RCurl")) {
        install.packages("RCurl")
}
if (!require("plyr")) {
        install.packages("plyr")
}
if (!require("tidyverse")) {
        install.packages("tidyverse")
}
if (!require("devtools")) {
        install.packages("devtools")
}
if (!require("narnia")) {
        install_github("njtierney/narnia")
}
if (!require("rnaturalearth")) {
        install.packages("rnaturalearth")
}

devtools::install_github("tidyverse/ggplot2")
library(ggplot2)
```

## Load data
```{r, message=FALSE,warning=FALSE,cache=TRUE, tidy=FALSE}
df <- readr::read_csv("https://goo.gl/F1tgk1")
```

## Data manipulation
```{r, message=FALSE,warning=FALSE,cache=FALSE}
# Convert names to lowercase
names(df) <- tolower(names(df))

# Rename variables
df <- plyr::rename(df,
                   c("merged id" = "merged_id",
                     "start date" = "start_date",
                     "end date" = "end_date",
                     "last name" = "last_name",
                     "first name" = "first_name",
                     "sex (c)" = "female",
                     "occupation (c)" = "occupation",
                     "occupation" = "occupation_2",                     
                     "victim affiliation (c)" = "victim_affiliation",
                     "victim affiliation" = "victim_affiliation2",
                     "violence (c)" = "violence",
                     "method (c)" = "method",
                     "interrogation (c)" = "interrogation",
                     "torture (c)" = "torture",
                     "war tribunal" = "war_tribunal",
                     "previous arrests (#)" = "number_previous_arrests",
                     "perpetrator affiliation (c)" = "perpetrator_affiliation",
                     "place-1" = "place_1",
                     "start location-1" = "start_location_1",
                     "latitude-1" = "latitude_1",
                     "longitude-1" = "longitude_1",
                     "exact coordinates-1" = "exact_coordinates_1",
                     "place-2" = "place_2",
                     "location-2" = "location_2",
                     "latitude-2" = "latitude_2",
                     "longitude-2" = "longitude_2",
                     "exact coordinates-2" = "exact_coordinates_2",
                     "place ©-3" = "place_3",
                     "end location-3" = "end_location_3",
                     "latitude-3" = "latitude_3",
                     "longitude-3" = "longitude_3",
                     "exact coordinates-3" = "exact_coordinates_3",
                     "place (c)-4" = "place_4",
                     "end location-4" = "end_location_4",
                     "latitude-4" = "latitude_4",
                     "longitude-4" = "longitude_4",
                     "exact coordinates-4" = "exact_coordinates_4",
                     "place (c)-5" = "place_5",
                     "end location-5" = "end_location_5",
                     "latitude-5" = "latitude_5",
                     "longitude-5" = "longitude_5",
                     "exact coordinates-5" = "exact_coordinates_5",
                     "place (c)-6" = "place_6",
                     "end location-6" = "end_location_6",
                     "latitude-6" = "latitude_6",
                     "longitude-6" = "longitude_6",
                     "exact coordinates-6" = "exact_coordinates_6"
                     ))

# Transform variables
df <- transform(df,
                merged_id = as.integer(merged_id),
                start_date = as.Date(start_date, "%d/%m/%Y"),
                end_date = as.Date(end_date, "%d/%m/%Y"),
                minor = as.factor(minor),
                age = as.numeric(age),
                female = female - 1,
                violence = as.integer(violence),
                method = as.integer(method),
                number_previous_arrests = as.integer(number_previous_arrests),
                perpetrator_affiliation = as.integer(perpetrator_affiliation),
                place_1 = as.integer(place_1),
                exact_coordinates_1 = as.numeric(exact_coordinates_1),
                exact_coordinates_2 = as.numeric(exact_coordinates_2),
                exact_coordinates_3 = as.numeric(exact_coordinates_3),
                exact_coordinates_4 = as.numeric(exact_coordinates_4),
                exact_coordinates_5 = as.numeric(exact_coordinates_5)
                )

# Recode torture and interrogation
df$torture <- ifelse(df$torture > 2 | is.na(df$torture), "NA",
                     ifelse(df$torture == 2, 0, df$torture))
df$interrogation <- ifelse(df$interrogation > 2 | is.na(df$interrogation), "NA",
                           ifelse(df$interrogation == 2, 0, df$interrogation))

# Revalue variables
df$minor <- plyr::revalue(df$minor, c("y" = 1, "n" = 0))

# Save data
readr::write_csv(df, "dataset.csv")
```

## Visualise Missing Data

```{r, warning=FALSE,error=FALSE,message=FALSE,fig.width=7, fig.height=7}
# Proportion of variables that contain any missing values
miss_var_pct(df)

# Number of missings in each variable
gg_miss_var(df)
```

\newpage

## Figure 1
```{r, warning=FALSE,error=FALSE,message=FALSE,cache=TRUE}
df1 <- df %>%
        filter(!is.na(violence)) %>%
        group_by(violence) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3)) %>%
        as.data.frame() %>%
        mutate(violence = factor(violence,
                                 labels = c("1" = "Killed",
                                            "2" = "Suicide",
                                            "3" = "Disappearance",
                                            "4" = "Disappearance\nwith Information",
                                            "5" = "Unresolved"))) %>%
        print()

ggplot(df1, aes(x = reorder(violence, -n), y = n)) +
        geom_histogram(stat = "identity") +
        scale_x_discrete("\nType of Violence") +
        geom_text(stat = "identity", aes(label = n),
                  vjust = -0.5, size = 2.6) +
        ylab("Number of Victims\n") +
        theme_minimal()  +
        theme(axis.text.y = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank())
```

\newpage

## Figure 2
```{r, warning=FALSE,error=FALSE,message=FALSE}
df2 <- df %>%
        select(age) %>%
        na.omit() %>%
        unlist() %>%
        as.numeric() %>%
        cut(breaks = c(0, 16, 20, 25, 30, 35, 40, 45, 50, 85)) %>%
        as.vector()

my_labels <- c("Under 16", "16-20", "20-25", "25-30", "30-35",
                "35-40", "40-45", "45-50", "Over 50")

ggplot(as.data.frame(df2), aes(x = df2)) +
        geom_histogram(stat = "count") +
        scale_x_discrete("", labels = my_labels) +
        geom_text(stat = "count", aes(label = ..count..),
                  vjust = -1, size = 2.6) +
        ylab("") +
        theme_minimal() +
        theme(axis.text.y = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank())
```

\newpage

## Figure 3
```{r, warning=FALSE,error=FALSE,message=FALSE, fig.width=8.75, fig.height=5}
df3 <- df %>%
        select(occupation) %>%
        filter(occupation %in% c(1:10)) %>%
        na.omit() %>%
        group_by(occupation) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3)) %>%
        as.data.frame() %>%
        mutate(occupation = factor(occupation,
                                   labels = c("White\nCollar", "Blue\nCollar",
                                              "School\nStudent",
                                              "University\nStudent", "Military",
                                              "Non-Military\nGovernment",
                                              "Unemployed", "Unknown", "Ex-Military",
                                              "Housewife")))

ggplot(df3, aes(x = reorder(occupation, -n), y = n)) +
        geom_histogram(stat = "identity") +
        scale_x_discrete("") +
        geom_text(stat = "identity", aes(label = n),
                  vjust = -0.5,  size = 2.6) +
        ylab("") +
        theme_minimal() +
        theme(axis.text.y = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.text.x = element_text(angle = 45, hjust = 1, size = 11))
```

\newpage

## Tables 1, 2 and 3
```{r}
df %>%
        select(interrogation) %>%
        group_by(interrogation) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3))

df %>%
        select(torture) %>%
        group_by(torture) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3))

df %>%
        select(targeted) %>%
        group_by(targeted) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3))

```

\newpage

## Figure 4
```{r, warning=FALSE,error=FALSE,message=FALSE, fig.width=8, fig.height=5.55}
df5 <- df

df5$female <- factor(df5$female,
                     levels = c(0, 1),
                     labels = c("Male", "Female"))

df5$violence <- factor(df5$violence,
                       levels = c(1, 2, 3, 4, 5),
                       labels = c("1" = "Killed",
                                  "2" = "Suicide",
                                  "3" = "Disappearance",
                                  "4" = "Disappearance\nwith Information",
                                  "5" = "Unresolved"))

par(mar = c(0, 0, 0, 0))
mosaicplot(female ~ violence, data = df5, las = 1, color = TRUE,
           xlab = "Gender", main = "", ylab = "Type of Violence")

table(df5$violence, df5$female)

round(prop.table(table(df5$violence, df5$female)), 2)
```

\newpage

## Figure 5
```{r, warning=FALSE,error=FALSE,message=FALSE}
df4 <- df %>%
        select(violence, start_date) %>%
        na.omit() %>%
        mutate(violence_dummy = ifelse(violence >= 1, 1, 0)) %>%
        mutate(day = format(start_date, "%d"),
               month = format(start_date, "%m"),
               year = format(start_date, "%Y")) %>%
        filter(year < 1992 & year > 1972) %>%
        group_by(year, month) %>%
        summarise(total = sum(violence_dummy)) %>%
        mutate(dates = paste(month, year, sep="-")) %>%
        mutate(day = "01", date = paste0(day, "-", dates),
               date = lubridate::parse_date_time(date, "dmY"))

ggplot(df4, aes(date, total)) +
        geom_line() +
        scale_x_datetime(name = "",
                         date_breaks = "15 months",
                         expand = c(0, 0),
                         date_labels = "%b\n%Y") +
        ylab("") +
        theme_minimal(base_size = 12) +
        theme(panel.grid.minor = element_blank(),
              axis.text.x = element_text(hjust = 1))
```

\newpage

## Figure 6
```{r}
df4 <- df %>%
        select(violence, start_date, occupation) %>%
        na.omit() %>%
        filter(occupation %in% c(1:6)) %>%
        mutate(violence_dummy = ifelse(violence >= 1, 1, 0)) %>%
        mutate(day = format(start_date, "%d"),
               month = format(start_date, "%m"),
               year = format(start_date, "%Y")) %>%
        filter(year < 1992 & year > 1972) %>%
        group_by(year, month, occupation) %>%
        summarise(total = sum(violence_dummy)) %>%
        mutate(dates = paste(month, year, sep="-")) %>%
        mutate(day = "01", date = paste0(day, "-", dates),
               date = lubridate::parse_date_time(date, "dmY")) %>%
        as.data.frame() %>%
        mutate(occupation = factor(occupation,
                                   labels = c("White\nCollar", "Blue\nCollar",
                                              "School\nStudent", "University\nStudent",
                                              "Military", "Non-Military\nGovernment")))

ggplot(df4, aes(date, total)) +
        geom_line() +
        scale_x_datetime(name = "",
                         date_breaks = "18 months",
                         expand = c(0, 0),
                         date_labels = "%b\n%Y") +
        ylab("") +
        theme_minimal(base_size = 12) +
        scale_y_continuous(expand = c(0, 0)) +
        theme(panel.grid.minor = element_blank(),
              axis.text.x = element_text(hjust = 1)) +
        facet_wrap( ~ occupation, scales = "free", ncol = 2)
```

\newpage

## Figure 7
```{r, warning=FALSE,error=FALSE,message=FALSE,cache=TRUE, fig.width=11,fig.height=7}
df7 <- df %>%
        select(violence, longitude_1, latitude_1) %>%
        na.omit()

df7$violence <- factor(df7$violence,
                       levels = c(1, 2, 3, 4, 5),
                       labels = c("1" = "Killed",
                                  "2" = "Suicide",
                                  "3" = "Disappearance",
                                  "4" = "Disappearance\nwith Information",
                                  "5" = "Unresolved"))

sf_data <- rnaturalearth::ne_countries(returnclass = "sf", continent = c("South America", "North America"))

ggplot() +
  geom_sf(data = sf_data, colour = "black", size = .2, fill = "#eeeeee") +
  theme_minimal() +
  geom_point(data = df7, aes(x = longitude_1, y = latitude_1,
                             color = factor(violence)),
             alpha = 0.4, size = 1) +
        theme(axis.title.x = element_blank()) +
        theme(axis.title.y = element_blank()) +
        theme(legend.position = "none") +
        facet_wrap(~ violence, ncol = 3) +
        theme(plot.margin = unit(c(0, 0, 0.5, 0), "cm"))
```

\newpage

## Figure 8
```{r, warning=FALSE,error=FALSE,message=FALSE,cache=TRUE,fig.width=11,fig.height=7}
df8 <- df %>%
        select(occupation, violence, longitude_1, latitude_1) %>%
        filter(occupation %in% c(1:10)) %>%
        na.omit() %>%
        group_by(occupation) %>%
        as.data.frame() %>%
        mutate(occupation = factor(occupation,
                                   labels = c("White\nCollar", "Blue\nCollar",
                                              "School\nStudent",
                                              "University\nStudent", "Military",
                                              "Non-Military\nGovernment",
                                              "Unemployed", "Unknown",
                                              "Ex-Military",
                                              "Housewife"))) %>%
        mutate(violence = ifelse(violence >= 1, 1, 0))

sf_data <- rnaturalearth::ne_countries(returnclass = "sf", continent = c("South America", "North America"))

ggplot() +
  geom_sf(data = sf_data, colour = "black", size = .2, fill = "#eeeeee") +
  theme_minimal() +
  geom_point(data = df8, aes(x = longitude_1, y = latitude_1,
                             color = factor(occupation)),
             alpha = 0.4, size = .8) +
        theme(axis.title.x = element_blank()) +
        theme(axis.title.y = element_blank()) +
        theme(legend.position = "none") +
        facet_wrap(~ occupation, ncol = 4) +
        theme(plot.margin = unit(c(0, 0, 0.5, 0), "cm"))
```

\newpage
```{r}
sessionInfo()
```

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{6pt}