### 2019-03-12 manuscript/article.Rmd
Add files via upload

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments. Data and replication information are available at [`http://github.com/danilofreire/pinochet-dataset`](http://github.com/danilofreire/pinochet-dataset)."
abstract: "... However, . This paper presents a georeferenced event dataset on "
keywords: "authoritarian regimes, Chile, geocoded event data, Pinochet, truth commission"
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: single
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    template: template.latex
---

# Introduction\label{sec:intro}

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

# Conclusion: New Avenues for Research

The dataset we introduce in this paper describes more than 2,000 human rights violations perpetrated by the Pinochet regime from 1973 to 1990. We include rich information about the events, such as the specific dates when the abuses occurred, personal details of the victims, and the geographical locations of the incidents. In that regard, our dataset provides a quantitative presentation of, and an extension to, the findings of Report of the Chilean National Commission on Truth and Reconciliation [-@report1991]. All information contained in the Report, plus our own spatial coordinates, are available in spreadsheet format which scholars can readily employ in their analyses.

Our data opens a number of potential avenues of research. For instance, @lupu2017legacy, @rozenas2017political and @zhukov2018stalin highlight that state repression has enduring effects on the political preferences and social attitudes of the victims. Researchers can test whether victims of the Pinochet regime show similar changes in political attitudes. Moreover, sociologists and criminologists may be interested in the relationship between the intensity of the human rights abuse and post-regime levels of social unrest or interpersonal violence.

The dataset also enables researchers to examine how political coalitions affect the use of lethal violence in authoritarian regimes. While the topic has a long tradition in political science [e.g., @de2010leader; @gandhi2007authoritarian], only recently authors have

The internal dynamics of autocratic governments remains an understudied topic in political science due to a lack of fine-grained information. But there are some noteworthy exceptions.

@fjelde2010generals and @rivera2017authoritarian are noteworthy exceptions. Both authors note that

By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime. XXXXXXXXXXXXX.

Finally, scholars can investigate the connections between international legitimacy and domestic politics in repressive regimes. This is a promising area of research as the Chilean government and American intelligence services continue to declassify documents from the Pinochet era. One relevant question is whether pressure from foreign governments and organisations had any influence over the levels of human rights abuses in Chile. We hope our dataset is useful for scholars interested in these and other questions, and that the information it contains elicits hypotheses not only about the Pinochet period, but about authoritarian governments more generally.

[^df]: Postdoctoral Research Fellow, The Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io). Corresponding author.

[^jm]: Senior Lecturer in Public Policy, Department of Political Economy, King's College London, [`john.meadowcroft@kcl.ac.uk`](mailto:john.meadowcroft@kcl.ac.uk), [`http://johnmeadowcroft.net`](http://johnmeadowcroft.net).

[^ds]: Associate Professor, Department of Political Science and the Political Theory Project, Brown University, [`david_skarbek@brown.edu`](david_skarbek@brown.edu), [`http://davidskarbek.com`](http://davidskarbek.com).

[^eg]: Software Developer, Tempo UK.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{6pt}