### 2021-03-01 manuscript/article.Rmd
add intro

---
title:  |
        | Human Rights Violations in the Pinochet Regime:
        | A Quantitative Presentation of the Rettig and Valech Reports
author: 
  - Mauricio Carrasco[^auth01]
  - Danilo Freire[^auth02]
  - Pablo Seguel Gutierrez[^auth03]
date: \today
thanks: "The authors thank Eugenia Guerrero, John Meadowcroft, and David Skarbek for their work on the Rettig Report dataset. All replication materials and the source code for the `pinochet R` package are available at <https://github.com/danilofreire/pinochet>."
abstract: "This article introduces a new dataset on human rights violations during the Pinochet regime in Chile (1973-1990). Based on previous work by @freire2019dataset, we code all cases of deaths and disappearances recorded in the Rettig (1991), Valech I (2004), and Valech II (2005) reports. Our dataset includes XX variables and presents information on XXXX individuals, of which XXX appear in at least two reports. We also provide geographical coordinates for all locations of human rights abused mentioned in the three reports. The data are available in several formats, including an `R` package built especifically for our purposes. The paper shows the temporal and spatial variation of human rights violations during the regime, and proposes further avenues of research for those interested in Chilean history and cross-national studies of state-sponsored violence."
abstractspacing: double
keywords: Chile; human rights violations; Pinochet; state-sponsored violence; truth commission
jelcodes: 
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)

# Chunk definitions
def_chunk_hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def_chunk_hook(x, options)
  ifelse(options$size != "normalsize",
         paste0("\n \\", options$size, "\n\n", x,
                "\n\n \\normalsize"), x)
})

# Install packages
packages <- c("kableExtra", "scales", "tidyverse")
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
invisible(lapply(packages, library, character.only = TRUE))

# Read dataset
library(readr)
df <- read_csv("../data/victims.csv")
```

[^auth01]: Author information. 
[^auth02]: Independent Researcher,
[`danilofreire@gmail.com`](mailto:danilofreire@gmail.com),
<https://danilofreire.github.io>.
[^auth03]: Author information.

\newpage

# Introduction
\label{sec:intro}

General Augusto Pinochet ruled Chile with an iron fist from 1973 to 1990. Later
findings show that the Chilean Armed Forces and the government's secret police
carried out gross human rights violations against political dissidents,
including murders, beatings, waterboarding, and kidnappings [@ensalaco2000chile;
@huneeus2007pinochet; @verdugo2001chile]. Most of the repression occurred in the
first months of the military regime, but state terrorism continued until the
last years of the dictatorship, often in collaboration with other Latin American
countries [@dinges2005condor; @mcsherry2012predatory]. While the true extension
of the brutality of the Pinochet regime remains unknown, estimates suggests that
the number of victims may be close to 30,000 [@weissbrodt1992report].

In this paper, we provide a quantitative presentation of all cases of deaths and
disappearances recorded in the Rettig and Valech reports. Both reports were
elaborated at the request of the Chilean government and are the official sources
of information on serious human rights abuses during the Pinochet years. The
National Commission for Truth and Reconciliation Report, also known as the
Rettig Report after its main editor Raúl Rettig, was compiled in 1991 and
recorded more than 2,000 cases of murder and kidnappings from 1973 to 1990
[@rettig1991report]. In 2004, a commission chaired by Bishop Sergio Valech
produced another compilation of victims of the military dictatorship, which
complemented and expanded the original findings [@valech2004report]. The Valech
report was updated in 2010 to include more cases.

Our dataset contains `r ncol(df)` variables and `r comma(nrow(df))`
observations. The data describes personal characteristics of the victims -- such
as unique id, name, age, gender, and occupation -- as well as information about
the crimes, including the type, perpetrator, and geolocation whenever such
information was available in the reports. We formatted the dataset following the
principles of "tidy data", in which each row correspondents to one case, each
variable forms a column, and the data are represented as a table
[@wickham2014tidy]. Moreover, we provide three different presentations of the
data. The first table includes only the cases mentioned in Rettig Report, the
second is the data from the Valech Report, and in the final table we merged all
cases together. To make the data easily available to users, we provide a
companion `R` package, named `pinochet` [@freire2019pinochet], alongside `csv`
and `xlsx` files that scholars can download at
<http://github.com/danilofreire/pinochet>.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}